---
title: "a3_task3_krista_finlay"
author: "Krista Finlay"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)
```

### Read in the urchin data

```{r}
urchin_data <- read_csv(here("data", "urchin_data.csv"))
```

## Section 1: Size comparisons of red and purple sea urchin
(include data from all sites, and all treatments)

### Exploratory data visualization: 
a.i. Comparing the distribution of sizes of red and purple urchins using the graph types- histogram and boxplot.
- Make a subset from the urchin_data set
- Clean column names and select columns size and common_name
-Then make ggplot histogram, separating the red and purple urchins using facet_wrap
-Label x and y axes



```{r}
urchin_size <- urchin_data %>% 
  clean_names() %>% 
  select(size, common_name)

ggplot(data = urchin_size, aes(x = size))+
  geom_histogram(binwidth = 0.5)+
  facet_wrap(~common_name)+
  labs(x = "Urchin Diameter (cm)", y = "Number of Urchins")

```
 
- Then using the urchin-size data set, make a ggpolt boxplot
- Label the x and y axes

```{r}
ggplot(data = urchin_size, aes(x = common_name, y = size))+
  geom_boxplot()+
  labs(x = "Common Name", y = "Urchin Diameter (cm)")

```

### Descriptive statistics table
a.ii. Creating a summary table that contains statistics for purple and red urchins:
- mean, median, standard deviation, sample size
- Use kableExtra to make the table look nicer


```{r}
urchin_summary <- urchin_size %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size), median = median(size), sd = sd(size),n = n())

urchin_summary

urchin_summary %>% 
  kable(col.names = c("Common Name", "Mean Diameter (cm)", "Median Diameter (cm)", "Standard Deviation", "Sample Size (n)")) %>% 
  kable_styling(full_width = FALSE)
```

### Written description regarding how the size (distribution, central tendancy, spread, etc) compare between the two species of urchins.

```{r}
#Actual Difference difference in mean urchin diameter between red and purple urchin species
6.776489 - 4.469797

# The % different in mean urchin diameter between red and purple urchin species
(6.776489- 4.469797)/(6.776489) *100
```
The purple urchins seem to have a smaller size distribution and not much variety in the diameters sizes. Purple sea urchin diameter sizes have a range that is smaller in comparison to red urchins, with the purple spreading mainly between 1 and 8 cm in diameter, with the bulk of the purple urchins found in the range of 3.5 to 5cm. The red urchins has a diameter size spread or range that is larger, as their diameter sizes can go up to 11cm, with a few even larger. Red urchins have more variability in their size distributions, compared to purple urchins. In comparison of the means between urchin species, the red urchins are 34% greater in size, which is a significant size difference, and may be associated with the distribution and range of sizes seen in the observations

## section 2: Exploring purple sea urchins by site and year

b.i. Create a subset of the urchins data specific for purple urchins:
- filter to only include observations of purple urchins
- Convert the date to class "Date"
- Create a new column that only contains the year
- Create a new column called site_full that contains the full name (not abbreviation) for each of the five sites: 
NAPL = Naples Reef, AQUE = Arroyo Quemado Reef, CARP = Carpinteria Reef, MOHK = Mohawk Reef, IVEE = Isla Vista Reef.

```{r}
clean_urchins <- urchin_data %>% 
  clean_names()

purple_urchins <- clean_urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(urchin_year = year(date_new)) %>% 
  mutate(site_full = case_when(site == "NAPL" ~"Naples Reef", site == "AQUE" ~"Arroyo Quemado Reef",
                               site == "CARP" ~"Carpinteria Reef", site == "MOHK" ~"Mohawk Reef",
                               site == "IVEE" ~"Isla Vista Reef"))
 
```


b.ii. Make a summary table:
- containing the total number of purple urchins observed by site and year
- starting from purple_urchins that was created above and find counts of observations by group. 


```{r}
purple_urchins_site_yr <- purple_urchins %>% 
  group_by(site_full, urchin_year) %>% 
  count(site_full)

purple_urchins_site_yr
```

b.iii. Make an exploratory graph of counts of purple urchins recorded at each site over time for the years included in the study.

```{r}
ggplot(data = purple_urchins_site_yr, aes(x = urchin_year, y = n, color = site_full))+
  geom_col()+
  facet_wrap(~site_full)+
  labs(x = "Year of Observation", y = "Number of Purple Urchins")+
  theme_gray()
```

